Verification of Stencil Computations through Spatial Specifications

### 2016-07-07 paper.tex

The sum of two regions $r \texttt{+} r'$ specifies that an array is
read using the neighbourhood indices described by either $r$ or
$r'$; it acts as a disjunction on specifications. For
example, the following gives the specification of a five-point
stencil which is the sum of two compound \texttt{reflexive} and
\texttt{centered} regions in each dimension:
%%
\begin{minted}{fortran}
!= stencil centered(depth=1, dim=1)*reflexive(dim=2) + centered(depth=1, dim=2)*reflexive(dim=1) :: a
b(i,j) = -4*a(i,j) + a(i-1,j) + a(i+1,j)
                   + a(i,j-1) + a(i,j+1)
\end{minted}
%%
Here the left-hand side of \texttt{+} says that when the second dimension
(induction variable $j$) is fixed at the origin, the first dimension
(induction variable $i$) accesses the immediate vicinity of the origin
(to depth of one). The right hand side of \texttt{+} is similar but the dimensions are reversed.
This reflects the symmetry under rotation of the five-point stencil.

As another example, the following
implements and specifies the \emph{Roberts cross}
edge-detection convolution~\cite{davis1975survey}:
%%
\begin{minted}[breakindent=5.5em]{fortran}
do j=0, jmax-1
    do i=0, imax-1
      x = a(i,j) - a(i+1,j+1)
      y = a(i+1,j) - a(i,j+1)
      != stencil forward(depth=1,dim=1) * forward(depth=1,dim=2) :: a
      b(i,j) = sqrt(x*x+y*y)
   end do
end do
\end{minted}

>>>>

The sum of two regions $r \texttt{+} r'$ specifies that an array is
read using the neighbourhood indices described by either $r$ or
$r'$; it acts as a disjunction on specifications. For
example, the following gives the specification of a five-point
stencil which is the sum of two compound \texttt{reflexive} and
\texttt{centered} regions in each dimension:
%%
\begin{minted}{fortran}
!= stencil centered(depth=1, dim=1)*reflexive(dim=2) + centered(depth=1, dim=2)*reflexive(dim=1) :: a
b(i,j) = -4*a(i,j) + a(i-1,j) + a(i+1,j) &
                   + a(i,j-1) + a(i,j+1)
\end{minted}
%%
Here the left-hand side of \texttt{+} says that when the second dimension
(induction variable $j$) is fixed at the origin, the first dimension
(induction variable $i$) accesses the immediate vicinity of the origin
(to depth of one). The right hand side of \texttt{+} is similar but the dimensions are reversed.
This reflects the symmetry under rotation of the five-point stencil.

As another example, the following
implements and specifies the \emph{Roberts cross}
edge-detection convolution~\cite{davis1975survey}:
%%
\begin{minted}[breakindent=5.5em]{fortran}
do j=0, jmax-1
  do i=0, imax-1
    x = a(i,j)-a(i+1,j+1); y = a(i+1,j)-a(i,j+1)
    != stencil forward(depth=1,dim=1) * forward(depth=1,dim=2) :: a
    b(i,j) = sqrt(x*x+y*y)
  end do
end do
\end{minted}
